Evaluating the evidence of widespread maintenance of functional composition in vertebrate communities

### 2021-10-07 paper/manuscript.Rmd
At table of model estimates.

knitr::opts_chunk$set(cache=FALSE, message=FALSE, warning=FALSE, out.width = "\\textwidth")

@@

```{r echo = FALSE, results = "asis"}
tibble(Taxa = c("Mammals", "Birds", "Amphibians"),
       Timeseries_Count = c(48, 2380, 11),
       Species_Count = c(184, 700, 184),
       Trait_Source = c("Elton Traits", "Elton Traits", "Amphibio"),
       Traits = c("body mass, diet, active diel period", "body mass, diet, nocturnality, \nforest foraging strata, pelagic specialist", "habitat, diet, active diel period, \nactivity seasonality, body mass, body length, \nmin maturation size, max maturation size, \nmin offspring size, max offspring size, \nreproductive output, breeding strategy")
) %>%
  mutate_all(linebreak, align = "l") %>% 
  kableExtra::kbl(format = "markdown", col.names = c("Taxa", "Number of \nTime Series", "Number of \nSpecies", "Trait Source", "Traits"), escape = FALSE, 
               caption = "Summary of the data in the final trait database.") #%>%
  #kable_classic() #%>% 
  #kable_styling(latex_options= c("scale_down", "hold_position")) #%>%
   #kableExtra::column_spec(2, width = "5cm") #%>%
  # kableExtra::column_spec(4, width = "10cm")
```

@@

We found no significant overall trend in species richness or functional diversity metrics (observed or standardized) (Fig 2).
There was also no temporal trend in those metrics within taxa, biome or realm factor levels.
We did find a significant overall decrease in Jaccard similarity, indicating increasing turnover through time.
Non-significant overall trends indicate that although some studies experience increasing or decreasing trends, the average trend across studies was plausibly 0.

At the study level, 4 studies experienced a significance trend in species richness and only 10 of 54 studies for observed metrics and 3 of 54 studies for standardized metrics experienced a significant trend for a metric other than Jaccard similarity (Table 2).
Most significant trends for observed functional metrics are in functional diversity and disappeared after standardization, indicating that richness increases were likely due to changes in the number of species.
<!-- Trends for all but one study were inconsistent between corrected and uncorrected metrics, with significance for one or the other. --> Hypothesis testing for study-level trends is likely affected by multiple testing issues and some trends identified as significant are therefore potentially erroneous.
Rather than interpreting changes in specific studies, we present these results as a general picture of the small number of studies experiencing a trend and highlight that even those studies are likely an overestimate of the number of significant trends.

@@

What does local maintenance of functional structure mean for the function of these communities?
The vast majority of experimental and observational work links declines in function to declines in functional or species diversity [@brose2016; @cadotte2011; @duffy2007] .
By those criteria, very few observed communities in our dataset are in a state of concern for loss of functionality.
However, we did not we did not directly measure some aspects of community structure important for ecosystem multifunctionality [e.g.
dispersion, rarity, abundance of dominant species; @bagousse-pinguet2021].
While our metrics indirectly incorporate dimensions of the functional space related to those community characteristics, they are likely not sensitive enough to infer deterioration (or improvement).
<!-- more citations please -->

>>>>

knitr::opts_chunk$set(cache=FALSE, message=FALSE, warning=FALSE, out.width = "\\textwidth")
options(knitr.kable.NA = '')

@@

```{r echo = FALSE, results = "asis"}
tibble(Taxa = c("Mammals", "Birds", "Amphibians"),
       Timeseries_Count = c(48, 2380, 11),
       Species_Count = c(184, 700, 184),
       Trait_Source = c("Elton Traits", "Elton Traits", "Amphibio"),
       Traits = c("body mass, diet, active diel period", "body mass, diet, nocturnality, \nforest foraging strata, pelagic specialist", "habitat, diet, active diel period, \nactivity seasonality, body mass, body length, \nmin maturation size, max maturation size, \nmin offspring size, max offspring size, \nreproductive output, breeding strategy")
) %>%
  mutate_all(linebreak, align = "l") %>% 
  kableExtra::kable(format = "latex", col.names = c("Taxa", "Number of \nTime Series", "Number of \nSpecies", "Trait Source", "Traits"), escape = FALSE, 
               caption = "Summary of the data in the final trait database.") %>%
  kable_classic() %>% 
  kable_styling(latex_options= c("scale_down", "hold_position")) #%>%
   #kableExtra::column_spec(2, width = "5cm") #%>%
  # kableExtra::column_spec(4, width = "10cm")
```

@@

We found no significant overall trend in species richness or functional diversity metrics (observed or standardized) (Fig 2).
There was also no temporal trend in those metrics within taxa, biome or realm factor levels.
We did find a significant overall decrease in Jaccard similarity, indicating increasing turnover through time.
Non-significant overall trends indicate that although some studies experience increasing or decreasing trends, the average trend across studies was plausibly 0 (Table 2).

```{r echo=FALSE}
read_csv(here::here("figures/man_model_table.csv")) %>%
  select(metric, everything()) %>%
  mutate(metric = ifelse(metric == "Jaccard_base", "Jaccard", metric)) %>%
  arrange(metric) %>%
  kable(format = "latex", caption = "Model estimates and statistics for general trend models for species richness (S) and standardized metrics. Additional model estimates can be found in the supplement.") %>%
  kable_classic() %>% 
  kable_styling(latex_options= c("scale_down", "hold_position"))
```

At the study level, 4 studies experienced a significance trend in species richness and only 10 of 54 studies for observed metrics and 3 of 54 studies for standardized metrics experienced a significant trend for a metric other than Jaccard similarity (Table 3).
Most significant trends for observed functional metrics are in functional richness and disappeared after standardization, indicating that richness increases were likely due to changes in the number of species.
<!-- Trends for all but one study were inconsistent between corrected and uncorrected metrics, with significance for one or the other. --> Hypothesis testing for study-level trends is likely affected by multiple testing issues and some trends identified as significant are therefore potentially erroneous.
Rather than interpreting changes in specific studies, we present these results as a general picture of the small number of studies experiencing a trend and highlight that even those studies are likely an overestimate of the number of significant trends.

@@

What does local maintenance of functional structure mean for the function of these communities?
The vast majority of experimental and observational work links declines in function to declines in functional or species diversity [@brose2016; @cadotte2011; @duffy2007] .
By those criteria, very few observed communities in our dataset are in a state of concern for loss of functionality.
However, we did not directly measure some aspects of community structure important for ecosystem multifunctionality [e.g.
dispersion, rarity, abundance of dominant species; @bagousse-pinguet2021].
While our metrics indirectly incorporate dimensions of the functional space related to those community characteristics, they are likely not sensitive enough to infer deterioration (or improvement).
<!-- more citations please --> This highlights the need for further development of theory linking different aspects of functional change to specific implications for function.
